The value of being indispensable: a cooperative approach to bargaining with a continuum of players

### 2023-11-29 src/paper/theory.tex
More or less finish example application

Following \textcite{armstrong2006competition}, let us also model player entry in a reduced form manner.
There is a continuum of potential entrants on both sides of the market.
Assume that the total utility of a player on side $i$ also includes an idiosyncratic component.
Let us order the players on side $i$ by this idiosyncratic utility in a decreasing manner.
That is, if we denote the idiosyncratic utility of player $n$ on side $i$ as $\epsilon_i(n)$, then $\epsilon_i$ is strictly decreasing in $n$.
Also assume that $\lim_{n \to \infty} \epsilon_i(n) = -\infty$.

@@

Due to the properties of $\epsilon_i$, this implies that there is a well-defined, finite number of entrants on each side.
Let us denote the number of entrants as a function of their non-idiosyncratic utility as $\phi_i(u_i)$.
Then, the total number of players on side $i$ is 
\begin{align*}
    n_i = \phi_i(u_i) = \epsilon_i^{-1}(u_i),
\end{align*}
where $\phi_i$ has the same role as in \textcite{armstrong2006competition}.

@@

One can also obtain the total welfare of side $i$, as it is simply the number of entrants plus the non-idiosyncratic utility of its players, plus the sum (integral) of the idiosyncratic utilities.
\begin{align*}
    w_i^*(n_i, n_j) = \alpha_i n_i n_j ^ {\gamma_i} + \gamma_j \alpha_j n_j n_i^{\gamma_j} - n_i F_i + \int_0^{n_i} \epsilon_i(s) \ds.
\end{align*}

@@

For ease of comparison with the bargaining case, let us also derive the utilities in this case:
\begin{align*}
    u_i^u(n_1, n_2) &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - \frac{\phi_1(u_i)}{\phi'_1(u_i)} \\
    &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - n_i \epsilon'_i(n_i).
\end{align*}
Welfare of side $i$ immediately follows in this case, too:
\begin{align*}
    U_i^u(n_1, n_2) = \alpha_i n_i n_j ^ {\gamma_i} + \gamma_j \alpha_j n_j n_i^{\gamma_j} - n_i F_i - n_i^2 \epsilon'_i(n_i) + \int_0^{n_i} \epsilon_i(s) \ds.
\end{align*}
Clearly, equilibrium utilities must be smaller than in the welfare-maximizing case, due to $\phi_i$ being increasing, the number of entrants is also lower.

@@

Finally, let us consider the case when the platform and the players bargain over the value they generate.
To do this, let us impose some structure on the timing of the game.
The idea is that even though the platform cannot commit to a specific entry fee structure anymore, it can commit to a bargaining process.
Specifically, it is able to commit to a bargaining weight for each side and itself: $\lambda_P, \lambda_1, \lambda_2$.

Then, the game proceeds as follows.
In the first period, the platform chooses $\lambda_P, \lambda_1$ and $\lambda_2$.
After that, potential antrants on both sides simultaneously decide whether to enter or not.
Finally, in the last period, the platform and the entrants bargain over the entry fees, and utilities are realized according to their weighted values.

Let us start by looking at the last period.
Assume that $n_1$ and $n_2$ players have decided to enter on sides $1$ and $2$, respectively.
Furthermore, assume that bargaining weights are $\lambda_1, \lambda_2$ and $\lambda_P$.

@@

The share received by each type of player is then easily established using \cref{prop:many_sided_weighted}.
\begin{proposition}
    \label{prop:platform_bargaining_last_period}
    Let $f(t_1, t_2) = w(t_1 n_1, t_2 n_2)$. Then the weighted values of the various sides are
    \begin{alignat*}{4}
        \pi_P^b(n_1, n_2) &= \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_P^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_P^{b_2}} \alpha_2 n_2 n_1^{\gamma_2} &&- \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_1}}_{\coloneqq S_P^{c_1}} n_1 F_1 &&- \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_2}}_{\coloneqq S_P^{c_2}} n_2 F_2 \\
        U_1^b(n_1, n_2) &= \underbracket{\frac{\lambda_1}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_1^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_1 \gamma_2}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_1^{b_2}}  \alpha_2 n_2 n_1^{\gamma_2} &&- \underbracket{\frac{\lambda_1}{\lambda_P + \lambda_1}}_{\coloneqq S_1^{c_1}} n_1 F_1 && \\
        U_2^b(n_1, n_2) &= \underbracket{\frac{\lambda_2\gamma_1}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_2^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_2}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_2^{b_1}} \alpha_2 n_2 n_1^{\gamma_2} && &&- \underbracket{\frac{\lambda_2}{\lambda_P + \lambda_2}}_{\coloneqq S_2^{c_2}} n_2 F_2.
    \end{alignat*}
\end{proposition}

@@

Let us look at the per-player utilities to

>>>>

Following \textcite{armstrong2006competition}, let us also model player entry in a reduced form manner.
There is a continuum of potential entrants on both sides of the market.
Assume that the total utility of an entrant on side $i$ also includes an idiosyncratic cost component.
Let us order the players on side $i$ by this idiosyncratic cost in an increasing manner.
That is, if we denote the idiosyncratic utility of player $n$ on side $i$ as $\kappa_i(n)$, then $\kappa_i$ is strictly increasing in $n$.
Also assume that $\lim_{n \to \infty} \kappa_i(n) = \infty$.

@@

Due to the properties of $\kappa_i$, this implies that there is a well-defined, finite number of entrants on each side.
Let us denote the number of entrants as a function of their non-idiosyncratic utility as $\phi_i(u_i)$.
Then, the total number of players on side $i$ is 
\begin{align*}
    n_i = \phi_i(u_i) = \kappa_i^{-1}(u_i),
\end{align*}
where $\phi_i$ has the same role as in \textcite{armstrong2006competition}.

@@

One can also obtain the total welfare of side $i$, as it is simply the number of entrants plus the non-idiosyncratic utility of its players, plus the sum (integral) of the idiosyncratic utilities.
\begin{align}
    \label{eq:W_star}
    W_i^*(n_i, n_j) = \alpha_i n_i n_j ^ {\gamma_i} + \gamma_j \alpha_j n_j n_i^{\gamma_j} - n_i F_i - \int_0^{n_i} \kappa_i(s) \ds.
\end{align}

@@

For ease of comparison with the bargaining case, let us also derive the utilities in this case:
\begin{align*}
    u_i^u(n_1, n_2) &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - \frac{\phi_1(u_i)}{\phi'_1(u_i)} \\
    &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - n_i \kappa'_i(n_i).
\end{align*}
Welfare of side $i$ immediately follows in this case, too:
\begin{align*}
    \label{eq:W_u}
    W_i^u(n_1, n_2) = \alpha_i n_i n_j ^ {\gamma_i} + \gamma_j \alpha_j n_j n_i^{\gamma_j} - n_i F_i - n_i^2 \kappa'_i(n_i) - \int_0^{n_i} \kappa_i(s) \ds.
\end{align*}
Clearly, equilibrium utilities must be smaller than in the welfare-maximizing case, due to $\phi_i$ being increasing, the number of entrants is also lower.

@@

Finally, let us consider the case when the platform and the players bargain over the value they generate.
I will approach this problem in two steps.
First, I will examine the distribution of the value generated in the case of no idiosyncratic cost.
I do this because this partial result has a neat interpretation, and is very tractable.
After that, I derive how the total idiosyncratic cost is distributed between the players.
Because of the linearity of random order values, the total welfare in the general case is simply the sum of these two building blocks.

Let us start by examining the case when there is no idiosyncratic cost.
Assume that $n_1$ and $n_2$ players have decided to enter on sides $1$ and $2$, respectively.
Furthermore, assume that bargaining weights are $\lambda_1, \lambda_2$ and $\lambda_P$.

@@

The share received by each type of player is then easily established using \cref{prop:many_sided_weighted}.
\begin{proposition}
    \label{prop:platform_bargaining_last_period}
    The weighted values of the various sides without idiosyncratic costs are
    \begin{alignat*}{4}
        \pi_P^b(n_1, n_2) &= \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_P^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_P^{b_2}} \alpha_2 n_2 n_1^{\gamma_2} &&- \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_1}}_{\coloneqq S_P^{c_1}} n_1 F_1 &&- \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_2}}_{\coloneqq S_P^{c_2}} n_2 F_2 \\
        w_1^b(n_1, n_2) &= \underbracket{\frac{\lambda_1}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_1^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_1 \gamma_2}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_1^{b_2}}  \alpha_2 n_2 n_1^{\gamma_2} &&- \underbracket{\frac{\lambda_1}{\lambda_P + \lambda_1}}_{\coloneqq S_1^{c_1}} n_1 F_1 && \\
        w_2^b(n_1, n_2) &= \underbracket{\frac{\lambda_2\gamma_1}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_2^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_2}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_2^{b_1}} \alpha_2 n_2 n_1^{\gamma_2} && &&- \underbracket{\frac{\lambda_2}{\lambda_P + \lambda_2}}_{\coloneqq S_2^{c_2}} n_2 F_2.
    \end{alignat*}
\end{proposition}

@@

Let us now turn our attention to the idiosyncratic cost component.
Assume that, even though entrants with different costs have different contribution to the total value, there can be no discrimination between them.
That is, they each must share the same fraction of the total idiosyncratic cost (e.g. have the same value).

Let us denote the total idiosyncratic cost incurred by all players on both sides as $K(n_1, n_2)$.
Also, suppose that if $n_i$ players enter on side $i$, then they are the ones with the lowest idiosyncratic cost.
Then, the total idiosyncratic cost is simply the sum of the idiosyncratic costs of the entrants on both sides:
\begin{align*}
    K(n_1, n_2) = \int_0^{n_1} \kappa_1(s) \ds + \int_0^{n_2} \kappa_2(s) \ds.
\end{align*}
Again, one can obtain the weighted values of all sides using \cref{prop:many_sided_weighted}.
\begin{proposition}
    \label{prop:platform_bargaining_idiosyncratic}
    The weighted values of the idiosyncratic cost component are
    \begin{align*}
        K_P(n_1, n_2) &= \sum_{i \in \{1, 2\}} \int_0^{n_1} \left[ 1 - \lambda_i \left( \frac{s}{n_i} \right)^{\lambda_P + \lambda_i - 1} \right] \kappa_i(s) \ds, \\
        K_i(n_1, n_2) &= \lambda_i \int_0^{n_1} \left( \frac{s}{n_i} \right)^{\lambda_P + \lambda_i - 1} \kappa_i(s) \ds.
    \end{align*}
\end{proposition}

That is, each side only has to bear some part of the total idiosyncratic cost, and the rest is shared with the platform.
As with the rest of the value, the share of the idiosyncratic cost is increasing in one's own bargaining weight.
This might seem counterintuitive at first, as one may expect that a player with a higher bargaining weight would be able to shift more of the cost to the other side.
However, in the case of random order values, this intuition is only correct for monotonic games, which it is not in this case.
One has to keep in mind though that this is just part of the final payoffs, and the other parts are increasing in one's own bargaining weight.
Therefore, as long as the combined game is monotonic, the final payoffs will be increasing in one's own bargaining weight.

Let us finally consider total welfare.
As mentioned before, it is simply the sum of the value generated and the idiosyncratic cost.
That is, the welfare of side $i \in \{1, 2\}$ is
\begin{align*}
    W_i^b(n_1, n_2) &= \frac{\lambda_i}{\lambda_P + \lambda_i + \lambda_j\gamma_i} \alpha_i n_i n_j^{\gamma_i} + \frac{\lambda_i \gamma_j}{\lambda_P + \lambda_j + \lambda_i\gamma_j} \alpha_j n_j n_i^{\gamma_j} - \frac{\lambda_i}{\lambda_P + \lambda_i} n_i F_i \\
    &- \lambda_i \int_0^{n_1} \left( \frac{s}{n_i} \right)^{\lambda_P + \lambda_i - 1} \kappa_i(s) \ds.
\end{align*}
It is remarkably similar in structure to the welfare maximizing case in \cref{eq:W_star}.
The main difference is that each part (both costs and positive utilities) is reweighted, and side $i$ only receives a fraction of it.
It is also worth noting that $W_i^b(n_1, n_2) < W_i^*(n_1, n_2)$, and, as a consequence, entry is lower than what would be socially optimal.\footnote{
    The easiest way to see it is that in the welfare-maximizing case the platform's profit is negative, while in the bargaining case it is positive.
    Therefore, fringe utilities must be lower in the latter case.
}

For better comparison, let us also examine implied entry fees and non-idiosyncratic utilities.
The latter is the difference between the average welfare and the average idiosyncratic cost:
\begin{align*}
    u_i^b(n_1, n_2) &= \frac{\lambda_i}{\lambda_P + \lambda_i + \lambda_j\gamma_i} \alpha_i n_j^{\gamma_i} + \frac{\lambda_i \gamma_j}{\lambda_P + \lambda_j + \lambda_i\gamma_j} \alpha_j n_j n_i^{\gamma_j - 1} - \frac{\lambda_i}{\lambda_P + \lambda_i} F_i \\
    &+ \int_0^{n_1} \left[ 1 - \lambda_i \left( \frac{s}{n_i} \right)^{\lambda_P + \lambda_i - 1} \right] \kappa_i(s) \ds.
\end{align*}
If one wished to close the model, the number of entrants in equilibrium could be obtained by solving the system of equations $\kappa_i(n_i) = u_i(n_1, n_2), n \in \{1, 2\}$.
Finally, the implied entry fees can be calculated from the equation $u_i = \alpha_i n_j^{\gamma_i} - p_i$, yielding
\begin{align*}
    p_i^b(n_1, n_2) &= \frac{\lambda_i}{\lambda_P + \lambda_i} F_i - \frac{\lambda_i \gamma_j}{\lambda_P + \lambda_j + \lambda_i\gamma_j} \alpha_j n_j n_i^{\gamma_j - 1} + \frac{\lambda_P + \lambda_j\gamma_i}{\lambda_P + \lambda_i + \lambda_j\gamma_i} \alpha_i n_j^{\gamma_i} \\
    &- \int_0^{n_1} \left[ 1 - \lambda_i \left( \frac{s}{n_i} \right)^{\lambda_P + \lambda_i - 1} \right] \kappa_i(s) \ds.
\end{align*}
